Using best practices for forecasting biodiversity changes in breeding birds

### 2017-07-13 manuscript/draft.Rmd
Add R and R package citations

We used observed richness values from 1982 (the first complete year of NDVI
data) to 2003 for training the models, and data from 2004 to 2013 was used for
testing their performance. We only used BBS routes from the continental United
States (i.e. routes where PRISM data was available), and restricted the analysis
to routes that were sampled during 70% of the years in the training period were
used (i.e., routes with at least 16 annual observations). The resulting dataset
included [[N]] transect runs of [[N]] unique sites, and included [[N]] species.
Site-level richness varied from [[N]] to [[N]] with an average richness of [[N]]
species.

@@

@@

In general, each model's error increased more-or-less linearly with time,
although the slopes and initial error rates varied across models (Figure 3). The
site-level models (i.e. the baselines and the ARIMA methods) tended to do very
well (Figure 3), especially on short time horizons, and especially in terms of
RMSE. Of these, the "Naive" model performed especially well at sites with
uncorrected observer turnover (since its predictions were best-calibrated to
observers that were active in later years).

Most models produced confidence intervals that were too narrow, indicating
overconfident predictions. The random forest-based SDM stack was the most
overconfident model, with observed richness values frequently falling well
outside its confidence limits (Figure 3C). This stacked SDM's narrow
predictive distribution also caused it to have notably higher deviance (Figure
3B) than the next-worst model, even though its point estimates were not
unusually bad in terms of RMSE (3A), and the other models' relative performance
was not nearly as sensitive to the choice of error metric. As discussed
elsewhere [@harris2015], this overconfidence is a product of the assumption in
stacked SDMs that errors in the species-level predictions are independent.
Because they avoided this assumption, the GBM-based "macroecological" model and
especially the mistnet JSDM had reasonably well-calibrated uncertainty
estimates, even in the presence of uncorrected observer biases (Figure 3C). This
calibration improved both models' deviance relative to the baseline models,
especially as the observer population evolved (Figure 3B). The "naive" model was
the only model whose confidence intervals were too wide (Figure 3), which can be
attributed to the rapid rate at which these intervals expand as the time lag
increases (Figure 1).

Accounting for the uncertainty associated with differences among observers
resulted in improvements to nearly all metrics of fit (Figure 4). The
improvements in point estimates (as measured by RMSE) primarily resulted from
large improvements in a small number of forecasts where a large difference
between observers resulted in a large shift in predicted richness values that
the models could not otherwise accommodate. The "Average" and ARIMA models
generally showed the largest improvement here, and the Naive model showed the
smallest (since it was already very responsive to observer turnover). As a
result, the "Average" and "ARIMA" models tended to perform the best, once
observer differences were accounted for. Improvements in coverage were notable
for all models except Naive and JSDM, meaning that unmodeled observer effects
had been partly responsible for the poor quantification of uncertainty. As a
result, the average and ARIMA models maintained a small average advantage over
the JSDM in terms of deviance at longer time scales, rather than falling behind
after about five years as in Figure 3B.

>>>>

We used observed richness values from 1982 (the first complete year of NDVI
data) to 2003 for training the models, and data from 2004 to 2013 was used for
testing their performance. We only used BBS routes from the continental United
States (i.e. routes where PRISM data was available), and restricted the analysis
to routes that were sampled during 70% of the years in the training period
(i.e., routes with at least 16 annual observations). The resulting dataset
included [[N]] transect runs of [[N]] unique sites, and included [[N]] species.
Site-level richness varied from [[N]] to [[N]] with an average richness of [[N]]
species.

@@

All analyses were conducted using R [@R]. Primary R packages used in the
analysis included dplyr [@dplyr], tidyr [@tidyr], gimms [@gimms], sp
[@sp1; @sp2], raster [@raster], prism [@prism], rdataretriever
[@rdataretriever], forecasts [@forecast1; @forecast2], git2r [@git2r], ggplot
[@ggplot2], viridius [@viridis], rstan [@rstan], yaml [@yaml], purrr [@purrr],
gbm [@gbm], randomForest [@randomForest], purrrlyr [@purrrlyr]. Code to fully
reproduce this analysis is available on GitHub
(https://github.com/weecology/bbs-forecasting) and archived on Zenodo ().

@@

While the models generally did well in absolute terms (Figure 2), they tended
not to outperform the "average" baseline from the generalized linear mixed model
(Figure 3A). [[Stats go here]].  The auto-ARIMA was generally the best-performing
non-baseline model, but in many cases, the best-fitting ARIMA model chosen by
the model selection procedure was no different from the "average" model.

Most models produced confidence intervals that were too narrow, indicating
overconfident predictions (Figure 3B). The random forest-based SDM stack was the
most overconfident model, with only XX% of observations falling inside its 95%
confidence intervals. This stacked SDM's narrow predictive distribution also
caused it to have notably higher deviance (Figure 3C) than the next-worst model,
even though its point estimates were not unusually bad in terms of RMSE (3A). As
discussed elsewhere [@harris2015], this overconfidence is a product of the
assumption in stacked SDMs that errors in the species-level predictions are
independent. Because they avoided this assumption, the GBM-based
"macroecological" model and especially the mistnet JSDM had reasonably
well-calibrated uncertainty estimates (Figure 3B); as a result, their relative
performance was higher in terms of deviance than in terms of RMSE. The "naive"
model was the only model whose confidence intervals were too wide (Figure 3B),
which can be attributed to the rapid rate at which these intervals expand as the
time lag increases (Figure 1).

In general, each model's error increased more-or-less linearly with time,
although the slopes and initial error rates varied across models (Figure 4). The
mistnet JSDM and GBM-based macroecological model's deviance tended to rise less
rapidly over time than the baseline models' deviance did; for the ten-year
forecasts, these environment-based models outperformed the Naive model and 
nearly matched the performance of the Average model.

When the differences among observers were not explicitly accounted for, nearly
all metrics of fit worsened (Figure 5). The increased error resulted from a
small number of forecasts where observer turnover caused a large shift in the
reported richness values. The Naive model was less sensitive to these shifts,
because it largely ignored the richness values reported by observers that had
retired by the end of the training period (Figure 1). The Average model, which
gave equal weight to observations from the whole training period, showed a
larger decline in performance -- especially in terms of coverage. The mistnet
JSDM was notable here because the coverage of its prediction intervals retained
very good coverage in the presence of this observer-based noise, which we
attribute to its ability to model this variation with its latent variables.
